### 2009-10-17 doc/tech_doc/janitor/tex_concept/concept.tex
Round of documentation updates - mostly semantics-invariant linguistic changes.

 The main language for the implementation of the functionality of the
 dynamic runtime environment functionality is Perl. And it is solely required (exceptions are the
 integration with the Grid Manager and the Web server) for the Janitor. In
 the pre-web-service implementation the Catalog remains a static web page.
 The Perl code is split into multiple modules as depicted in Figure~\ref{fig:janitorDependencies}. 
 The modules can be separated into two
 functional groups. One addresses the retrieval of information from
 the Catalog RDF file in the left major branch of the figure. The other
 addresses the process of fetching and installing the packages.

@@

\section{Runtime environments states}
A main motivation for the managed, manual initiation of dynamic RE
installation is the subsequent manual verification of the installed
packages -- prior to their use in production.
With an automation of the installation, the verification of
that process shall be performed externally to that process.  At this
time, only the automation of the installation has been implemented.
To reflect the progress the external verification has made, REs
are said to be in states. The current implementation lists
installable REs aside the installed REs in the grid information system,
in order to stimulate grid clients to submit packages. The here described
states will be represented to the clients in upcoming developments.

These states are specific for every CE and communicated
between the Janitor and the Execution Service. Table~\ref{tab:states} shows all
possible states, while Figure~\ref{fig:RE_states}
displays the transitions between the states that a Runtime Environment may
be in during its life time at a particular CE. \task{?at a particular CE?}

@@

@@

The manually induced transitions are marked in red, he automated transitions in black.
A transition between states can be induced automatically (i.\,e.\ by the
advent of a job requesting a particular dynamic RE) or manually by the
site's supervisor or an individual with respective rights to use the
Janitor's web service.

@@

% Runtime Environment are installed for job idenitified by a certain id.

@@

Janitor has two states for jobs: \texttt{PREPARED} and \texttt{INITIALIZED}. After a job has been succesfully registered 
in Janitor, its state will be set to \texttt{PREPARED}. Invalid jobs are not cached in Janitor. 
After Janitor is requested to deploy the runtime environment, the state of the job will change to \texttt{INITIALIZED}. If 
an unforeseen exception occures during that process, Janitor will drop the job from its database and set the affected runtime
environments to the state \texttt{FAILED}.

@@

\newenvironment{note}
{\rule{1ex}{1ex}\hspace{\stretch{1}}}
{\hspace{\stretch{1}}\rule{1ex}{1ex}\\}
\begin{note}
The integration into AREX is not completed yet!
\end{note}
\begin{note}
Thus, there will be bigger changes here in this section\dots
\end{note}

@@

Currently the Janitor perl scripts must be executed as root. In order to execute Janitor with the required permissions,
a setuid wrapper has been written. During the installation \task{not done yet} the mode of rjanitor will be set to \texttt{u+s}
to make the binary suid root. \task{According to Daniel: ``A future version of the Janitor will get rid of the suid root helper.''}\\

Thus, Janitor is not executed directly. Instead a dynamic link will be created in \task{libexec or sbin} which is pointing to
the wrapper.\\

It is suggested to create a user and a group "janitor" for the Janitor.\\

% Currently the Janitor must be executed as root. For this a suid root helper
% is needed. It is provided in /opt/janitor/rjanitor.c. If you installed
% the Janitor in another directory then /opt/janitor change the parameter
% of the execv command in rjanitor.c. Then compile the file with\\
% \texttt{\textdollar gcc -o rjanitor rjanitor.c}\\
% and run as root\\
% \texttt{\textdollar chown root:root rjanitor}\\
% \texttt{\textdollar chmod u+s rjanitor}\\
% to make the binary suid root. A future version of the Janitor will get rid
% of the suid root helper.\\
% 
% The Grid Manager does not call the Janitor directly in /opt/janitor but
% calls /opt/nordugrid/libexec/janitor. So a link is needed (as root):\\
% \texttt{\textdollar ln -s /opt/janitor/rjanitor /opt/nordugrid/libexec/janitor}\\
% \\
% It is suggested to create a user and a group "janitor" for the Janitor.\\
% \\
% The Janitor needs a directory for storing information on installed REs:\\
% \texttt{\textdollar mkdir -p /var/lib/nordugrid/janitor}\\
% \texttt{\textdollar chown janitor:janitor /var/lib/nordugrid/janitor}\\
% \\
% It also needs a directory on the shared volume for installing the REs:\\
% % \texttt{\textdollar mkdir -p $<$shared_volume$>$/janitor}\\
% % \texttt{\textdollar chown janitor:janitor $<$shared_volume$>$/janitor}

@@

Malevolent regular users with respective training in using system exploits
to gain root access are likely to find security holes by regularly
submitted scripts.  The authentication and authentification of users,
together with respective logging, is the major defense against such
attacks. What is consequently left to be protected against are unwanted
sideeffects by the installation of software.

The worst case scenario would be the installation of a RE
that overwrites system files. With the current implementation, which is
based solely on tar files, this is barely possible, unless such is
performed by the install scripts that accompany the tar files. However,
hereto the installation would have to be performed by a user with system
priviledges, for which there is no technical requirement.

@@

Summarising, there is general concern about the security of grid
computing.  Dynamic REs introduce new dangers since a
manual control at the grid site is substituted by a remote process that
is out the direct supervision of a local site administrator. The signing
of packages by known and directly or indirectly trusted developers is
a good indicator that no malevolent individuals have tampered with the
binary. The site administrators can limit the sources of packages and
specify packages that are eligible or excluded from installations.

>>>>

The main language for the implementation of the functionality of the
dynamic runtime environment functionality is Perl. And it is solely required (exceptions are the
integration with the Grid Manager and the Web service) for the Janitor. In
the pre-web-service implementation the Catalog remains a static web page.
The Perl code is split into multiple modules as depicted in Figure~\ref{fig:janitorDependencies}. 
The modules can be separated into two
functional groups. One addresses the retrieval of information from
the Catalog RDF file in the left major branch of the figure. The other
addresses the process of fetching and installing the packages.

@@

\section{Runtime environments states}

A main motivation for the managed, manual initiation of dynamic RE
installation is the subsequent manual verification of the installed
packages -- prior to their use in production.

With an automation of the installation, the verification of
that process shall be performed externally to that process.  At this
time, only the automation of the installation has been implemented.
To reflect the progress the external verification has made, REs
are said to be in states. The current implementation lists
installable REs aside the installed REs in the grid information system,
in order to stimulate grid clients to submit packages. The here described
states will be represented to the clients in upcoming developments.

These states are specific for every CE and communicated
between the Janitor and the execution service \AREX. Table~\ref{tab:states} shows all
possible states, while Figure~\ref{fig:RE_states}
displays the transitions between the states that a Runtime Environment may
be in during its life time.

@@

The concept of ARC prevails to have a single directory into which to
install software to be execute on all compute elements of the site.
Otherwise, the installation of a runtime environment would be
required to be performed just when the jobs initiates its computation.

In principle this would be doable, particularly for those runtime 
environments that are part of a trusted Linux distribution already.
But this would also mean that the compute nodes become inhomogeneous,
and at the time of writing this was considered undesireable.

@@

The manually induced transitions are marked in red, he automated transitions in black.
A transition between states can be induced automatically (\ie by the
advent of a job requesting a particular dynamic RE) or manually by the
site's administrator or an individual with respective rights to use the
Janitor's command line.

@@

\subsection{Dependencies on Job IDs}

The Janitor was designed with the job execution in mind. As such, all
actions it performs are driven by the need of a job. And those jobs
should then be specified, to learn when the demand for a particular
runtime environment has ended.

This concept partially conflicts with the idea to use the Janitor
as an aid for the manual invocation of an installation -- there is no
job ID that could be assigned for that process. Today, the administrator
is suggested to use a special number, \eg 0, to indicate such manual
installs. This will change in near future.

@@

The Janitor has two states for jobs: \texttt{PREPARED} and \texttt{INITIALIZED}.
After a job has been succesfully registered with the Janitor, its state
will be set to \texttt{PREPARED}. Invalid jobs are not cached by the Janitor. 
After the Janitor is requested to deploy the runtime environment, the
state of the job will change to \texttt{INITIALIZED}. If an unforeseen
exception occures during that process, Janitor will drop the job from
its database and set the affected runtime environments to the state
\texttt{FAILED}.

@@

\newenvironment{note}
{\rule{1ex}{1ex}\hspace{\stretch{1}}}
{\hspace{\stretch{1}}\rule{1ex}{1ex}\\}

@@

At the time of writing, the Janitor needs to be executed as root. 
This may change over time, but as long as all other components
of ARC also demand superuser privileges, one can expect the
demand for the Janitor to be non-problematic.

@@

Malevolent regular users with respective training in using system exploits
to gain root access are likely to find security holes by regularly
submitted scripts.  The authentication and authentification of users,
together with respective logging, is the major defense against such
attacks. What is consequently left to be protected against are unwanted
side-effects by the installation of software.

The worst case scenario would be the installation of a RE
that overwrites system files. With the current implementation, which is
based solely on tar files, this is barely possible, unless such is
performed by the install scripts that accompany the tar files.
%However,
%hereto the installation would have to be performed by a user with system
%priviledges, for which there is no technical requirement.

@@

As a conclusion, there is general concern about the security of grid
computing.  Dynamic REs introduce new dangers since a manual control
at the grid site is substituted by a remote process that
is out the direct supervision of a local site administrator. The signing
of packages by known and directly or indirectly trusted developers is
a good indicator that no malevolent individuals have tampered with the
binary. The site administrators can limit the sources of packages and
specify packages that are eligible or excluded from installations.

### 2009-10-17 doc/tech_doc/janitor/tex_future/future.tex
Round of documentation updates - mostly semantics-invariant linguistic changes.

\section{Representation of dynamic REs in the information model}
Dynamic REs require an extended representation in the
information model.  The Application Software description should be able
to distinguish installed REs from installable REs, potentially offer
description of extended RE state-like information. This work is planned
to be carried out as part of the Glue-2.0 effort of OGF\footnote{OGF
GLUE: https://forge.gridforum.org/sf/projects/glue-wg}.

@@

The RDF schema nicely prepares for the upcoming virtualisation of worker
nodes. Hereto, the {\tt BaseSystem} indicates a virtual image to which
further packages, the dynamic REs, would then be added.
How exactly the dynamics are integrated will depend on how
dynamic the virtualisation of the nodes is. In the simplest scenario,
a worker node's CPU will only be occupied by a single virtual machine
and that will not be changed. In this case, there is no difference to
the setup of the Janitor with today's static setups.

@@

TODO:\\
\begin{itemize}
 \item There is a arc.conf file in which all possiblie flags are listed... ADD THE JANITOR FLAGS!!!
 \item Dynamic RTEs are now listed as manually installed in the ``janitor list'' command.. change that
 \item Verfication
\end{itemize}

%(1) The HostInfoJanitor.pm queries the catalog for available RTE names.
%    It is fast, such that I assume a fork or a cache is currently not needed. 
%    > Nevertheless I installed an alarm of 15 Seconds

%(2) In order to enable the Grid-Manager to invoke Janitor, a wrapper rJanitor.c has
%    (already) been written (by Daniel). 
%    This wrapper gives Janitor the necessary rights using a suid bit. 
%    >It is installed and used but not yet tested well
%    >In case Daniel is right, it will work

% (3) > ... ?? hmm... well seems there is some information loss here... point 3... suppose it must have been important.

%(4) The function "check" (nowadays ``info'') was ment to return the information about a runtime 
%      environment. It may possible to merge with the deploy functionality 
%      if desired. (The Janitor source code will remain the same, only the interface will 
%      be changed.)
%      >Due to this is beautification, this is postponed

%(5) The information about the runtime environments is passed to the backend
%    script. (It is not specifed yet how this has to be done)
%       o Installation directories
%       o Script to set environment variables
%    > Currently a ``deploy'' action will set symbolic links for *ALL* dynamically installed
%    > METAPACKAGES!  That means, nothing has to be done here. The backends script won't (hopefully)
%    > except the symbolic links and execute the corresponding script delievered along with the tarball 
%    > package. 
%    > The symbolic link will be removed respictvely on ``remove''
%
%    > TODO: Dynamic RTEs are now listed as manually installed in the ``janitor list'' command.. change *that*
%
%    > TODO: Think about the user rights of the symbolic link...

%(6) Once the grid job has been finished, the token on the runtime environment will
%      be removed. (If no tokens left, the environment may be deinstalled)
%    > Well, that's nothing new for me, but maybe nice to know :-)

%(7) GLUE2 specification: Balazs like to have an interface for Janitor. This task is 
%   put on the agenda. It is still unclear how the interaction with that interface shall be 
%   realised. Janitor can only access data concerning the class ApplicationEnvironment.
%   Shall a XML snippet be returned?
%  > Alright now, almost 1.5 month later I know what he asked for.
%  >
%  > Please examine following files in src/services/a-rex/infoproviders/:
%    o ARC1ClusterInfo.pm:~744
%    o HostInfo.pm:~218
%    o HostInfoJanitor.pm
%  >
%  > The HostInfo.pm collects a list of manually and dynamically (HostInfoJanitor.pm) installed RTE *NAMES*.
%  > ARC1ClusterInfo.pm adds some hard coded information (description, states, etc.) to it and puts it into the schema.
%  >
%  > WHAT HAS TO BE DONE?
%  >
%  > HostInfo.pm has to create a *HASH MAP* with must have as much information as possible (description, states, etc.).
%  > Therefore HostInfoJanitor.pm has to be extended to gather more information out of the catalog.
%  > ARC1ClusterInfo.pm needs to be stultified such that the *HASH MAP* received from HostInfo.pm will simply mapped
%  > to the schema.

%(8) Verfication
%    o Of validity of tar-package (Store the catalog somewhere safe! Hash-value database?)
%    o Of the executer of Janitor (This can be done using a WEB SERVICE!!! )
%    o TODO: I experienced some problems in case some shell tools are not installed on clusters
%            i.e.  unzip
%            There should be some kind of check or RESTRICTION to what can be used

%(9) The missing feature for dependencies 
%    Assume, one wants to install package A which depends on package B.
%    Both are getting installed dynamically.
%    Right before job execution, the runtime scripts of package A is executed. 
%    Still missing is the execution of the runtime script of B!
%
%    Simple workaround: Request package B in job description too
%    Correct solution:  Dependencies have to be returned by Janitor and added to the job.description file!!!!!
%                       (Janitor info 21 or *EXTEND* Janitor deploy + Get Janitor output in states.cpp and put
%			it into the description file)

>>>>

\section{Representation of dynamic REs in the information model}
Dynamic REs require an extended representation in the
information model.  The Application Software description should be able
to distinguish installed REs from installable REs, potentially offer
descriptions of extended RE state-like information. This work is planned
to be carried out as part of the Glue-2.0 effort of OGF\footnote{OGF
GLUE: https://forge.gridforum.org/sf/projects/glue-wg}.

@@

The RDF schema nicely prepares for the upcoming virtualisation of worker
nodes.  How exactly the dynamics are integrated will depend on how
dynamic the virtualisation of the nodes is. In the simplest scenario,
a worker node's CPU will only be occupied by a single virtual machine
and that will not be changed. In this case, there is no difference to
the setup of the Janitor with today's static setups.

@@

\section{Use of RDF}

The RDF is "correctly" used within the Janitor via the Redland libraries.
What is still missing is a semantical reasoning on what packages to
allow or disallow. For instance: allow all packages associated with
bioinformatics. This shall wait a bit longer until more experiences
on the community's demand for this service have been made.

\section{Manual Verification}

There is yet no explicit notion of a concept on how users can report
on the reliability of individual compute nodes and use such information
for the decision making on where to sent their jobs. And with an increasing
complexity of software installed, being used across versions, one will
rather trust one's very own experiences for individual sites than some
external repository or other pieces of information that may be weeks
or months old.

Consequently, no means for a manual verification and the communication
of results of such have yet been implemented.

\section{Known issues}
\begin{itemize}
 \item There is a arc.conf file in which all possiblie flags are listed... ADD THE JANITOR FLAGS!!!
 \item Dynamic RTEs are now listed as manually installed in the ``janitor list'' command.. change that
\end{itemize}

### 2009-10-17 doc/tech_doc/janitor/tex_installation/installation.tex
Round of documentation updates - mostly semantics-invariant linguistic changes.

% ** Dependencies
% *** Ordinary: log4perl and ``Redland RDF Library - Perl Interface''
% *** WebService: libperl-dev
% *** Protege: For maintaining the knowledge base

The Janitor requires the two perl packages listed in
table~\ref{tab:install_dependencies}. To have
the WebService interface for the Janitor, the packages listed in the
table~\ref{tab:install_dependencies_optional} need to be installed before
the build process is. The Perl modules are available on CPAN and ship
with all major Linux distributions.

\begin{table}[!h]
   \begin{center}
        \mycaption{Required perl packages for Janitor.}{Log4perl is used for the internal logging of Janitor, while the 
                   Redland RDF library is used for accessing the knowledge base of Runtime Environments.}
        \label{tab:install_dependencies}
	\begin{tabular}{|p{3cm}|p{7cm}|}
	\hline
	   liblog-log4perl-perl & \textit{Log4perl is a port of the log4j logging package}\\
	\hline
	   librdf-perl          & \textit{Perl language bindings for the Redland RDF library}\\
	\hline
	\end{tabular} 
   \end{center}
\end{table}
\begin{table}[!h]
   \begin{center}
        \mycaption{Optional libraries for Janitor.}{The library libperl-dev provides header files which are needed to link the 
                   WebService to the Perl interpreter.}
        \label{tab:install_dependencies_optional}
	\begin{tabular}{|p{3cm}|p{7cm}|}
	\hline
	   libperl-dev & \textit{ Perl library: development files}\\
	\hline
	\end{tabular}
   \end{center}
\end{table}
\forcelinebreak

The Janitor will be installed by default along with ARC1. Support for
the dynamic REs can be completely disabled at compile time by using
the \textit{configure} flags \textit{--disable-janitor-service} for
the complete janitor or \textit{--disable-janitor-webservice} to only
disable the Janitor's Web Service.

Furthermore it is recommended to install the ontology editor 
Prot\`eg\`e\footnote{\href{http://protege.stanford.edu}{http://protege.stanford.edu}} in order to be to easily maintain the 
knowledge database of installable packages. At the time of writing, no Linux distribution is offering packages for this fine tool.

@@

The current version of the Janitor can be configured using the common file \textit{arc.conf}. It is expected in the 
configuration directory \textit{etc}. The Janitor is using the environment variable \textit{NORDUGRID\_CONFIG}
to determine the location of the corresponding file. If that variable is not set, the default location \textit{/etc/arc.conf} will be used. 
The configuration is assigned by the section \lbrack janitor\rbrack. Table~\ref{tab:arc_conf_tags} describes the available tags
for Janitor's configuration.

\begin{landscape}
\begin{table}[!h]
   \begin{center}
        \mycaption{Tags usable in \textit{arc.conf} within the section janitor.}{Tags usable in \textit{arc.conf} within the section janitor.}
	\label{tab:arc_conf_tags}
	\begin{longtable}{|p{3cm}|p{10cm}||p{10cm}|}
	\hline
	   \textbf{tag}    & \textbf{example}                      & \textbf{description}\\
        \hline
           enabled         & "1"                                   & Boolean flag which enables or disables janitor in A-REX.\\
	   uid             & "root"                                & The effective uid. \\
	   gid             & "0"                                   & The effective gid. \\
	   registrationdir & "/var/spool/nordugrid/janitor"        & Directory where we the current states of jobs are kept. \\
	   catalog         & "/var/spool/nordugrid/janitor/catalog/knowarc.rdf"& URL of the catalog containg the package information.\\
	   downloaddir     & "/var/spool/nordugrid/janitor/download" & Directory for downloads \\
	   installationdir & "/var/spool/nordugrid/janitor/runtime"& Directory for installation of packages                   \\
	   jobexpirytime   & "7200"                                & If a job is older than this, it is considered dead and assigned to be removal pending.\\
	   rteexpirytime   & "36"                                  & If a runtime environment was not used for this time, it will be assigned to be removal pending.\\
	   allow\_base     & "*"                                   & Allow rule for base packages. \\
	   deny\_base      & "debian::etch"                        & Deny rule for base packages.\\
	   allow\_rte      & "*"                                   & Allow rule for base packages. \\
	   deny\_rte       & "APPS/MATH/ELMER-5.0.2"               & Deny rule for base packages. \\
	   logconf         & "/opt/nordugrid/etc/log.conf"         & Location of the logging configuration file for janitor.\\
	\hline
	\end{longtable}
   \end{center}
\end{table}
\end{landscape}

The parameter \texttt{enable} specifies whether the Janitor shall be used within A-REX or not. Use the value \texttt{"0"} to disable Janitor. The
\texttt{uid} and the \texttt{gid} are defining which effective the user and group id shall be used for Janitor. 
The \texttt{registrationdir} describes the directory in which the subdirectories \texttt{jobs} and \texttt{rtes} will be created.
In these directories the states of the jobs and the runtime environments will be stored. 
The knowledge base of installable packages is specified by the parameter \texttt{catalog}. 
Its value can be any kind of URL pointing to a file written in the Resource Description Framework (RDF) format.
The specification of the RDF file will be explained in detail in section~\ref{sec:catalog}.
The parameter \texttt{downloaddir} assigns the directory in which the installation files will be saved after they have been 
downloaded or copied from the repository which was specified by the catalog. The \texttt{installationdir} finally specifies the
directory into which all packages will be installed and should be 
available for all computing elements, i.e.\ by using a shared volume.
If the configuration file furthermore contains the \texttt{runtimedir} tag within the section \texttt{grid-manager}, the Janitor will also
create a symbolic link in the \texttt{runtimedir} pointing to the configuration script of the installation done by the Janitor.
The tags \texttt{jobexpirytime} and \texttt{rteexpirytime} are used for automated cleanup and is defined in seconds.
The default value for the \texttt{jobexpirytime} is seven days and for the \texttt{rteexpirytime} three days. 
The additional tags \texttt{allow\_base} \texttt{deny\_base} \texttt{allow\_rte} and \texttt{deny\_rte} are used to include
or exclude certain base packages or runtime environments of the catalog. This feature is useful, if the catalog is maintained by
a higher organization.
The path to the log4perl configuration file is defined by the tag \texttt{logconf}.
An example how to configure ARC and log4perl is provided in the Listings~\ref{lst:arc_conf} and ~\ref{lst:log_conf}.

@@

The Janitor is designed for the usage on Linux distributions.
Its porting to other platforms has not yet been addressed.

The ARC middleware is not ultimately essential for dynamic Runtime Environments.
All the Perl code would be functional with any Grid middleware.

% Is designed for linux distributions. Windows or Mac won't work (Reason wget! maybe also the Executer).

>>>>

% ** Dependencies
% *** Ordianary: log4perl and ``Redland RDF Library - Perl Interface''
% *** WebService: libperl-dev
% *** Protege: For maintaining the knowledge base

The Janitor depends on two perl packages listed in table~\ref{tab:install_dependencies}. To have the WebService interface for Janitor,
the packages listed in the table~\ref{tab:install_dependencies_optional} as build requirements.
\begin{table}[!h]
   \begin{center}
        \mycaption{Required perl packages for the Janitor.}{Log4perl is used for the internal logging of Janitor, while the 
                   Redland RDF libraray is used for accessing the knowledge base of RTEs.}
        \label{tab:install_dependencies}
	\begin{tabular}{|p{3cm}|p{7cm}|}
	\hline
	   liblog-log4perl-perl & \textit{Log4perl is a port of the log4j logging package}\\
	\hline
	   librdf-perl          & \textit{Perl language bindings for the Redland RDF library}\\
	\hline
	\end{tabular} 
   \end{center}
\end{table}
\begin{table}[!h]
   \begin{center}
        \mycaption{Optional libraries for Janitor.}{The library libperl-dev provides header files which are needed to link the 
                   WebService to the Perl interpreter.}
        \label{tab:install_dependencies_optional}
	\begin{tabular}{|p{3cm}|p{7cm}|}
	\hline
	   libperl-dev & \textit{ Perl library: development files}\\
	\hline
	\end{tabular}
   \end{center}
\end{table}
\forcelinebreak

If you are using regular Debian or Ubuntu packages, then the Janitor
can be installed as root by "apt-get install nordugrid-arc1-janitor".
Installing it will not drag other components of ARC with it since the
Janitor can be used in its own right or in conjunction with another
grid system, possibly. Packages for Redhat/Fedora and SuSE/OpenSuSE are
also provided, the redland library however may not yet be available for
those systems.  If you are compiling the sources yourself, the Janitor
will be enabled by default. If desired, it is possible to disable it
with the \textit{configure} flags \textit{--disable-janitor-service}
for the complete janitor or \textit{--disable-janitor-webservice} for
only the Web Service support.

Furthermore it is recommended to install the ontology editor 
Prot\`eg\`e\footnote{\href{http://protege.stanford.edu}{http://protege.stanford.edu}} in order to be to easily maintain the 
knowledge database of installable packages.

@@

The current version of Janitor can be configured using the common file \textit{arc.conf} which is to be found in the 
configuration directory \textit{etc}. Janitor is using the environment variable \textit{NORDUGRID\_CONFIG} to determine the location 
of the corresponding file. If the variable is not set, the default location \textit{/etc/arc.conf} will be used. 
The configuration is assigned by the section \lbrack janitor\rbrack. The table~\ref{tab:arc_conf_tags} contains the available tags
for janitor configuration.
\begin{landscape}
\begin{table}[!h]
   \begin{center}
        \mycaption{Tags usable in \textit{arc.conf} within the section janitor.}{Tags usable in \textit{arc.conf} within the section janitor.}
	\label{tab:arc_conf_tags}
	\begin{longtable}{|p{3cm}|p{10cm}||p{10cm}|}
	\hline
	   \textbf{tag}    & \textbf{example}                      & \textbf{description}\\
        \hline
           enabled         & "1"                                   & Boolean flag which enables or disables janitor in A-REX.\\
	   uid             & "root"                                & The effective uid. \\
	   gid             & "0"                                   & The effective gid. \\
	   registrationdir & "/var/spool/nordugrid/janitor"        & Directory where we the current states of jobs are kept. \\
	   catalog         & "/var/spool/nordugrid/janitor/catalog/knowarc.rdf"& URL of the catalog containg the package information.\\
	   downloaddir     & "/var/spool/nordugrid/janitor/download" & Directory for downloads \\
	   installationdir & "/var/spool/nordugrid/janitor/runtime"& Directory for installation of packages                   \\
	   jobexpirytime   & "7200"                                & If a job is older than this, it is considered dead and assigned to be removal pending.\\
	   rteexpirytime   & "36"                                  & If a runtime environment was not used for this time, it will be assigned to be removal pending.\\
	   allow\_base     & "*"                                   & Allow rule for base packages. \\
	   deny\_base      & "debian::etch"                        & Deny rule for base packages.\\
	   allow\_rte      & "*"                                   & Allow rule for base packages. \\
	   deny\_rte       & "APPS/MATH/ELMER-5.0.2"               & Deny rule for base packages. \\
	   logconf         & "/opt/nordugrid/etc/log.conf"         & Location of the logging configuration file for janitor.\\
	\hline
	\end{longtable}
   \end{center}
\end{table}
\end{landscape}

The parameter \texttt{enable} defines whether the Janitor shall be
used within \AREX or not. Use the value \texttt{"0"} to disable the
Janitor. The \texttt{uid} and the \texttt{gid} are defining which
effective user and group id shall be used for Janitor. Setting these to
values different from root will increase security.

The \texttt{registrationdir} describes the directory in which
the subdirectories \texttt{jobs} and \texttt{rtes} will be created.
In these directories the states of the jobs and the runtime environments
are stored. Please recall that the Janitor does not use a database as
a backend, but all communication between invocations are performed via
files in those folders.

The knowledge base of installable packages is specified by the parmeter
\texttt{catalog}.  Its value can be any kind of URL pointing to a file
written in the Resource Description Framework (RDF) format. One should
not light-heartedly use a remote address for this purpose. Such a remote
source needs to be trusted.

The specification of the RDF file will be explained in detail in
section~\ref{sec:catalog}.  The parameter \texttt{downloaddir} assignes
the directory in which the installation files will be saved after they
have been downloaded or copied from the repository which was specified by
the catalog. Please remember: the URL in arc.conf indicates the location
of the catalog. And the URLs somehow specified in the catalog specify
the location from where to download the runtime environment.

The \texttt{installationdir} finally specifies the directory into which
all packages will be installed. This \texttt{installationdir} directory
needs to be available for all computing nodes for the execution of 
arbitrary programs, mostl likely by using it as a shared NFS volume.

If the configuration file furthermore contains the \texttt{runtimedir}
tag within the section \texttt{grid-manager}, the Janitor will
also create a symbolic link in the \texttt{runtimedir} pointing to
the configuration script of the installation done by Janitor. The
tags \texttt{jobexpirytime} and \texttt{rteexpirytime} are used
for automated cleanup and are defined in seconds.  The default
value for thew\texttt{jobexpirytime} is seven days and for
the \texttt{rteexpirytime} three days.  The additional tags
\texttt{allow\_base}, \texttt{deny\_base}, \texttt{allow\_rte} and
\texttt{deny\_rte} are used to include or exclude certain base packages
or runtime environments of the catalog. This feature is useful, if the
catalog is maintain by a higher organization. But again: you need to
trust it.

The path to the log4perl configuration file is defined by the tag
\texttt{logconf}.  Examples how to configure arc and log4perl are
provided in the Listings~\ref{lst:arc_conf} and ~\ref{lst:log_conf}.

@@

The Janitor is only available for UNIX-based servers. While not tested,
it should also be functional on MacOS X and Windows with Cygwin.

### 2009-10-17 doc/tech_doc/janitor/tex_introduction/introduction.tex
Round of documentation updates - mostly semantics-invariant linguistic changes.

Janitor is a service for automated runtime environment installation for grid computing elements. It is integrated in the A-REX service used for job submission but also can be used as a standalone tool.

@@

A major motivation for grid projects is to stimulate new communities to adopt computational grids for their
causes. From the current grid user's viewpoint, the admission of users of a very different education will
suddenly impose difficulties in the communication between site maintainers. One will not even understand
the respective other side's research aims. Hence, the proper installation of non-standard software (Runtime
Environments) is not guaranteed and the priorities of manual labour will be mostly disjunctive.

A core problem remains to distribute a locally working solution, the Know-How, quickly across all contributing sites, i. e., without manual interference. Every scientific discipline has its respective own set of
technologies for the distribution of work load. E. g. research in bioinformatics requires access to so many
different tools and databases, that few sites, if any, install them all. Instead, the use of web services became
a commodity, with all the problems with respect to bottlenecks and restrictions of repeated access.
The EU project KnowARC1\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}} amongst other challenges with the here presented work extends the NorduGrid?s
Advanced Research Connector (ARC) grid middleware~\cite{ELLERT_2007} towards an infrastructure for the
automated installation of software packages.

@@

This document will start with a chapter of how to set up janitor locally. The following chapter will give an instruction on how to 
use Janitor with A-REX and/or without A-REX. Afterwards, in the third chapter, the maintenance of the program will be presented, 
which is basically covering the method how to prepare new runtime environments. A deeper knowledge about the design of Janitor will 
be given by the subsequent forth chapter. In the fifth chapter, an outlook to future developments will given.

>>>>

The {\em Janitor} is a service for the automated installation of runtime environments for grid computing elements.
Its command line interface allows for a direct interaction with site administrators. However, the main stimulus
for its development was the idea integrate such a service with the regular handling of compute jobs. For ARC this
is performed by the \AREX module.

From the programmer's view, the Janitor is mostly a Perl script and the routines within \AREX to invoke it. The
site administrator will also associate with it also the Catalog files, that describe the availability of runtime
environments, and the repository of installable runtime envionments themselves, which are regular tar archives obeying
to particular structure and reside in a separate folder. In order to minimise the latency for the invocation of
the Perl script and the associated parsing of files, a Janitor web service was developed, which still is a Perl
script.

@@

A major motivation for grid projects is to stimulate new communities to adopt computational grids for their
causes. From the current grid user's viewpoint, the admission of users of a very different scientific disciplines
or compute skills will impose difficulties in the communication between site maintainers. One will not even understand
the respective other side's research aims. Hence, the proper installation of non-standard software (read Runtime
Environments) is not guaranteed.

A core problem remains to distribute a locally working solution, the Know-How, quickly across all contributing
sites, i. e., without manual interference. Every scientific discipline has its respective own set of
technologies for the distribution of work load. For instance, research in bioinformatics requires access to so many
different tools and databases, that few sites, if any, install them all. Instead, the use of web services became
a commodity, with all the problems with respect to bottlenecks and restrictions of repeated access.
The EU project KnowARC\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}},  amongst other challenges,
with the here presented work extends the NorduGrid?s
Advanced Research Connector (ARC) grid middleware~\cite{ELLERT_2007} towards an the
automated installation of software packages.

@@

This document will start with a chapter of how to set up janitor locally. The following chapter will give an instruction on how to 
use Janitor with \AREX and/or without \AREX. Afterwards, in the third chapter, the maintenance of the program will be presented, 
which is basically covering the preparation of runtime environments. Deeper knowledge about the design of the Janitor will 
be given by the subsequent forth chapter. In the fifth chapter, an outlook on ongoing or future developments will presented.
